---
title: "Clinical Outcomes Causal Inference Report"
author: "Emily Qiu"
format:
  html:
    toc: true
    number-sections: true
execute:
  warning: false
  message: false
---

## Executive Summary

This report estimates whether a non-randomized clinical treatment is associated with lower 30-day readmission and improved event-free survival. Because treatment assignment is confounded by baseline severity, comorbidity, prior utilization, and hospital site, the analysis uses propensity score matching, inverse probability weighting, adjusted regression, and survival modeling.

The project is designed as a reproducible R workflow rather than a one-off notebook. The same functions are used by the `targets` pipeline, tests, report, and Shiny app.

## Setup

```{r}
library(dplyr)
library(ggplot2)
library(broom)
library(survival)
library(MatchIt)

source("../R/01_generate_data.R")
source("../R/02_clean_data.R")
source("../R/03_balance_checks.R")
source("../R/04_propensity_score.R")
source("../R/05_survival_analysis.R")
source("../R/06_outcome_models.R")
source("../R/07_visualizations.R")
```

## Data

```{r}
raw_data <- load_or_generate_clinical_data("../data/synthetic_clinical_outcomes.csv")
clinical_data <- clean_clinical_outcomes(raw_data)

dplyr::glimpse(clinical_data)
```

## Baseline Cohort Summary

```{r}
clinical_data |>
  group_by(treatment) |>
  summarise(
    patients = n(),
    mean_age = mean(age),
    mean_severity = mean(baseline_severity),
    mean_comorbidity = mean(comorbidity_index),
    readmission_rate = mean(readmitted_30d),
    event_rate = mean(event_observed),
    .groups = "drop"
  )
```

## Propensity Score Adjustment

```{r}
propensity_model <- fit_propensity_model(clinical_data)
scored_data <- add_propensity_scores(clinical_data, propensity_model)
matched_data <- create_matched_cohort(scored_data)
weighted_data <- create_iptw_cohort(scored_data)

unadjusted_balance <- calculate_balance_table(scored_data, "treatment")
matched_balance <- calculate_balance_table(matched_data, "treatment")
weighted_balance <- calculate_weighted_balance_table(weighted_data, "treatment", "iptw")

plot_balance_comparison(unadjusted_balance, matched_balance, weighted_balance)
```

## Propensity Score Overlap

```{r}
plot_propensity_overlap(scored_data)
```

## Binary Outcome Models

```{r}
outcome_models <- fit_outcome_models(clinical_data, matched_data, weighted_data)
tidy_outcome_results(outcome_models)
```

Interpretation: estimates below 1.0 indicate lower odds of 30-day readmission in the treated group relative to control.

## Survival Analysis

```{r}
survival_models <- fit_survival_models(clinical_data, matched_data, weighted_data)
tidy_survival_results(survival_models)
```

Interpretation: hazard ratios below 1.0 indicate lower event hazard in the treated group.

## Kaplan-Meier Curve

```{r}
plot_kaplan_meier(clinical_data)
```

## Limitations

- The dataset is synthetic and designed for portfolio demonstration.
- Causal claims depend on no unmeasured confounding.
- Propensity score methods improve balance on observed covariates only.
- Small samples can produce unstable matched estimates.
- A production study would require preregistered outcomes, clinical review, sensitivity analysis, and external validation.
